There and back again: a retrosynthesis tale

### 2019-03-30 papers/retrosynthesis/article.tex
Gathering data from individual models.

For this study, we implemented The Transformer model in Tensorflow library to support its integration in our in-house programs set (\url{https://github.com/bigchem/retrosynthesis}).  All values reported are averages for three repetitive runs. Preliminary modeling showed that the architecture with 3 layers and 8 attention works well for the datasets, though we tried combinations of 2, 6, 8, 10 layers with 6, 8, 10, 12 heads. So all calculations were performed with these values fixed.  The number of learnable parameters of the model is 1\,882\,176, embedding layer common for product and reactants has size 64.
\begin{table}[b!]
  \caption{Accuracy (\%) of the models on test set when all reactants were correctly predicted.}
  \centering
  \begin{tabular}{p{2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{8cm}}
    \toprule
    Model & top-1 & top-3 & top-5 & Description\\
    \midrule
    Seq2Seq & 37.4  & 52.4 & 57.0 & Literature result from \cite{Pande} based on Seq2Seq architecture. \\
    T1     & 38.27 & 58.23 & 64.96 & Transformer Model trained with validation control set (early stopping, 200 epochs). \\
    T$1_1$ & 40.16 & 58.95 & 65.54 & The same as T1, but without early stopping (1000 epochs). \\ 
    T2     & 42.30 & 62.29 & 68.95 & Transformer Model trained on both training and validation sets for 1000 epochs.\\
    T$2_1$ & 40.36 & & & T = 1.3 during the decoding stage of the T2 model. \\
    T3     &  & & & Transformer Model trained with cyclic learning rate schedule for 6000 epochs.\\
    T$3_1$ & 42.28  & & & T =  1.3 during the decoding stage of T3 model \\
    T4     &  42.25  & 61.31 & 68.11 & Transformer Model trained with distillation approach \cite{Hinton} for 1000 epochs.\\ 
    T$4_1$ & 42.91 & 62.11 & 68.92 & The same as T4 but decoding at T=1.3 \\
    T$4_1$ & 44.42 & 63.66 & 70.38 & The same as T4 but decoding at T=1.3, rdkit version 2017.09.3\\    \bottomrule
  \end{tabular}
  \label{tbl:results}
\end{table}

@@

\subsection{T1 model} 
T1 model was trained for 1000 epochs with training and validation datasets. Warmup steps for the learning rate is 16\,000, then it decreases.

\subsubsection{Model 1 (run-100)}

@@

\begin{table}[h!]
\caption{Parameters of T11 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Early stopping loss & 0.10864 & & & \\
    Last training loss & 0.11819 & & & \\
    Epoch (early stopping) & 208 & & & \\
    \midrule
    Greedy search (early stopping, epoch 208) & & 32.83 & &\\
    Greedy search (last, epoch 1000) & & 34.75 & & \\
    Greedy (last 10 average, epoch, 991 - 1000) & & 35.57 & & \\
    \midrule
    Beam, T = 1.0 (early) & & 35.57 & 54.14 & 59.13 \\
    Beam, T = 1.3 (early) & & 35.93 & 54.59 & 59.51 \\ 
    \midrule
    Beam, T = 1.0 (last) & &  &  &  \\
    Beam, T = 1.3 (last) & \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & &  &  &  \\
    Beam, T = 1.3 (last 10 average) & \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

@@

\begin{table}[h!]
\caption{Parameters of T21 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Last training loss & 0.05108 & & & \\
    \midrule
    Greedy search (last, epoch 1000) & & 38.85 & & \\
    Greedy (last 10 average, epoch, 991 - 1000) & &  & & \\
    \midrule
    Beam, T = 1.0 (last) & & 40.57 &60.19  & 66.77 \\
    Beam, T = 1.3 (last) & & 41.17 & 60.65 &  67.28 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & &  &  &  \\
    Beam, T = 1.3 (last 10 average) & &  &  &  \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

@@

\begin{table}[h!]
\caption{Parameters of T22 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Last training loss & 0.04943 & & & \\
    \midrule
    Greedy search (last, epoch 996) & &  & & \\
    Greedy (last 10 average, epoch, 991 - 1000) & &  & & \\
    \midrule
    Beam, T = 1.0 (last) & &  &  &  \\
    Beam, T = 1.3 (last) & &  &  &   \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & &  &  &  \\
    Beam, T = 1.3 (last 10 average) & &  &  &  \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

>>>>

For this study, we implemented The Transformer model in Tensorflow library to support its integration in our in-house programs set (\url{https://github.com/bigchem/retrosynthesis}).  All values reported are averages for three repetitive runs. Preliminary modeling showed that the architecture with 3 layers and 8 attention works well for the datasets, though we tried combinations of 2, 6, 8, 10 layers with 6, 8, 10, 12 heads. So all calculations were performed with these values fixed.  The number of learnable parameters of the model is 1\,882\,176, embedding layer common for product and reactants has size 64.
\begin{table}[b!]
  \caption{Accuracy (\%) of the models on test set when all reactants were correctly predicted.}
  \centering
  \begin{tabular}{p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}p{8cm}}
    \toprule
    Model & Greedy & Top-1 & Top-3 & Top-5 & Description\\
    \midrule
    Seq2Seq &  & 37.4  & 52.4 & 57.0 & Literature result from \cite{Pande} based on Seq2Seq architecture. \\
    T1         & &  &  &  & Transformer Model trained with validation control set (early stopping, 200 epochs). \\
    T$1_1$ & &  &  &  & The same as T1, but without early stopping (1000 epochs). \\ 
    T2         & 39.3 & 41.1 & 60.5 & 66.6 & Transformer Model trained on both training and validation sets for 1000 epochs.\\
    T$2_1$ & & 41.8 & 61.3 & 67.2 & T = 1.3 during the decoding stage of the T2 model. \\
    T3         & &  & & & Transformer Model trained with cyclic learning rate schedule for 6000 epochs.\\
    T$3_1$ & &   & & & T =  1.3 during the decoding stage of T3 model \\
    T4         & &  &  &  & Transformer Model trained with distillation approach \cite{Hinton} for 1000 epochs.\\ 
    T$4_1$ & &  &  &  & The same as T4 but decoding at T=1.3 \\
     \bottomrule
  \end{tabular}
  \label{tbl:results}
\end{table}

@@

\subsection{T1 model} 
\subsubsection{Model 1 (run-100)}

@@

\begin{table}[h!]
\caption{Parameters of T11 run.}
  \centering
  \begin{tabular}{p{6.2cm}p{3.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Early stopping loss (training / validation) &  0.09345 / 0.10864 & & & \\
    Last training loss (training / validation) &  0.05867 / 0.11819& & & \\
    Epoch (early stopping) & 208 & & & \\
    \midrule
    Greedy search (early stopping, epoch 208) & & 32.83 & &\\
    Greedy search (last, epoch 999) & & 34.75 & & \\
    Greedy (last 10 average, epoch, 990 - 999) & & 35.57 & & \\
    \midrule
    Beam, T = 1.0 (early) & & 35.57 & 54.14 & 59.13 \\
    Beam, T = 1.3 (early) & & 35.93 & 54.59 & 59.51 \\ 
    \midrule
    Beam, T = 1.0 (last) & & 37.15 & 55.22 &  60.49 \\
    Beam, T = 1.3 (last) & & 37.67 & 55.97 & 61.01 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & & 37.83 & 56.08 &  60.63 \\
    \textbf{Beam, T = 1.3 (last 10 average)}  & & \textbf{38.15} & \textbf{56.71} & \textbf{61.31} \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-1-early.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-1-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-1-avg.h5}

\subsubsection{Model 1 (run-101)}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 16.5cm]{images/t12.pdf}
  \caption{Learning curves and rate schedule for T1 model. Losses for training and validation datasets on the left, usual learning rate schedule on the right.}
  \label{fig:t11}
\end{figure}

\begin{table}[h!]
\caption{Parameters of T12 run.}
  \centering
  \begin{tabular}{p{6.2cm}p{3.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Early stopping loss (training / validation) &  0.08681 / 0.09926 & & & \\
    Last training loss (training / validation) &  0.04863 / 0.11316  & & & \\
    Epoch (early stopping) & 162 & & & \\
    \midrule
    Greedy search (early stopping, epoch 162) &  & 34.63 & &\\
    Greedy search (last, epoch 990) &  & 38.07  & & \\
    Greedy (last 10 average, epoch, 990 - 999) & & 38.19 & & \\
    \midrule
    Beam, T = 1.0 (early) & & 37.45 & 56.71 & 62.43 \\
    Beam, T = 1.3 (early) & & 38.31 & 57.93 & 63.35 \\ 
    \midrule
    Beam, T = 1.0 (last) & & 39.83 & 58.63 &  64.45 \\
    Beam, T = 1.3 (last) & & 40.67 &  59.53 & 64.97 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & & 40.22 & 59.05 &  64.91 \\
    \textbf{Beam, T = 1.3 (last 10 average)} & & \textbf{40.85} &  \textbf{59.85} & \textbf{65.45} \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t12}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-2-early.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-2-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-2-avg.h5}

\subsubsection{Model 1 (run-102)}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 16.5cm]{images/t1-3.pdf}
  \caption{Learning curves and rate schedule for T1 model. Losses for training and validation datasets on the left, usual learning rate schedule on the right.}
  \label{fig:t11}
\end{figure}

\begin{table}[h!]
\caption{Parameters of T13 run.}
  \centering
  \begin{tabular}{p{6.2cm}p{3.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Early stopping loss (training / validation) &  0.07795 / 0.09662 & & & \\
    Last training loss (training / validation) &  0.04654 / 0.11285  & & & \\
    Epoch (early stopping) & 193 & & & \\
    \midrule
    Greedy search (early stopping, epoch 193) &  & 35.63 & &\\
    Greedy search (last, epoch 998) &  & 38.13 & & \\
    Greedy (last 10 average, epoch, 990 - 999) & & 38.03 & & \\
    \midrule
    Beam, T = 1.0 (early) & & 38.75 & 58.55 &  64.47 \\
    Beam, T = 1.3 (early) & & 39.53 & 59.25 & 65.23 \\ 
    \midrule
    Beam, T = 1.0 (last) & & 40.05 & 59.35 &  64.27 \\
    Beam, T = 1.3 (last) & & 40.41 & 60.49 &  64.85 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & &  &  &   \\
    \textbf{Beam, T = 1.3 (last 10 average)} & &  &  &   \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t12}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-3-early.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-3-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t1-3-avg.h5}

@@

\begin{table}[h!]
\caption{Parameters of T21 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Last training loss & 0.05108 & & & \\
    \midrule
    Greedy search (last, epoch 981) & & 38.85 & & \\
    Greedy (last 10 average, epoch, 990 - 999) & & 39.25 & & \\
    \midrule
    Beam, T = 1.0 (last) & & 40.57 &60.19  & 66.77 \\
    Beam, T = 1.3 (last) & & 41.17 & 60.65 &  67.28 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & & 40.98 & 60.29 & 67.00 \\
    \textbf{Beam, T = 1.3 (last 10 average)} & &  \textbf{41.61} & \textbf{60.85} & \textbf{67.37} \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-1-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-1-avg.h5}

@@

\begin{table}[h!]
\caption{Parameters of T22 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Last training loss & 0.04943 & & & \\
    \midrule
    Greedy search (last, epoch 996) & & 38.79 & & \\
    Greedy (last 10 average, epoch, 990 - 999) & & 39.11 & & \\
    \midrule
    Beam, T = 1.0 (last) & & 40.98 & 60.39 & 66.38  \\
    Beam, T = 1.3 (last) & & 41.58 &  61.21 & 66.94 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & & 41.14  & 60.59 &  66.41\\
    \textbf{Beam, T = 1.3 (last 10 average)} & &  \textbf{41.87} &  \textbf{61.35} &  \textbf{67.08}\\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

@@

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-2-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-2-avg.h5}

\newpage
 \subsubsection{Model 2 (run-202)}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 16.5cm]{images/t2-3.pdf}
  \caption{Learning curves and rate schedule for T2 model. Loss for training on the left, usual learning rate schedule on the right.}
  \label{fig:t21}
\end{figure}

\begin{table}[h!]
\caption{Parameters of T23 run.}
  \centering
  \begin{tabular}{p{8.2cm}p{1.5cm}p{1.5cm}p{1.5cm}p{1.5cm}}
    \toprule
    Parameter & Value & Top-1 & Top-3 & Top-5 \\
    \midrule
    Last training loss & 0.04826 & & & \\
    \midrule
    Greedy search (last, epoch 984) & & 39.79 & & \\
    Greedy (last 10 average, epoch, 990 -  999) & & 39.69 & & \\
    \midrule
    Beam, T = 1.0 (last) & & 41.55 & 60.79 & 66.57  \\
    Beam, T = 1.3 (last) & & 41.85 & 61.63  & 67.32 \\ 
    \midrule
    Beam, T = 1.0 (last 10 average) & & 41.45  & 60.75 &  66.55\\
    \textbf{Beam, T = 1.3 (last 10 average)} & & \textbf{42.02}  & \textbf{61.73}  & \textbf{67.19} \\ 
    \bottomrule
  \end{tabular}
  \label{tbl:t11}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-3-last.h5}

\url{https://github.com/bigchem/retrosynthesis/papers/retrosynthesis/models/t2-3-avg.h5}
